rough

### 2021-10-20 old/algo2-elec16.Rmd
cleaning folder

parcoord <- GGally::ggparcoord(data_pcp,
                   columns = 2:ncol(data_pcp),
                   groupColumn = 1,
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_viridis(discrete=TRUE) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord

>>>>

parcoord <- GGally::ggparcoord(data_pcp,
                   columns = 2:ncol(data_pcp),
                   groupColumn = 1,
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax", 
                   order = "anyClass"
) + ggplot2::theme_bw() +
  scale_color_viridis(discrete=TRUE) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord

### 2021-10-20 old/interact_gran_4D.Rmd
cleaning folder

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("R/simdata01.R")
```

@@

```{r sim-val, echo = TRUE}
set.seed(9999)
nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 3 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 1 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
```

@@

```{r nobs50, echo = TRUE}
ntimes_val = 50 # nobs per combination 
```

@@

@@

```{r, echo=TRUE}
sample_seed <-  seq(1, 100, 1)
```

@@

```{r data-make}
data_null <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_null_normal) %>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varf_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varx_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varall_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

@@

>>>>

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
library(here)
read_chunk(here::here("old/simdata01.R"))
```

@@

```{r sim-val, echo = TRUE}
set.seed(9999)
nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 1 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 1 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
```

@@

```{r nobs50, echo = TRUE}
ntimes_val = 1000 # nobs per combination 
```

@@

p_null/p_varf/p_varx/p_varall
```

@@

```{r, echo=TRUE}
sample_seed <-  seq(1, 5, 1)
```

@@

```{r data-make}
data_null <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_null_normal) 
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varf_normal)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varx_normal)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varall_normal)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

@@

```{r}
data_clust <- data_q %>% tsibble::as_tsibble(key = unique_data, index = time)

library(gracsr)
data_clust %>% scale_gran(method="nqt", response = "sim_data")

%>%
  dist_gran(gran1="id_facet", gran2 = "id_x") %>% 
  dist_gran(clust_gran(kopt=4)

### 2021-10-20 old/rough.Rmd
cleaning folder

---
title: Clustering of residential Australian smart meter customers

# to produce blinded version set to 1
blinded: 0

authors:
- name: Sayani Gupta
  affiliation: Department of Econometrics and Business Statistics, Monash University
  thanks: "Email: Sayani.Gupta@monash.edu"

- name: Rob J Hyndman
  affiliation: Department of Econometrics and Business Statistics, Monash University

- name: Dianne Cook
  affiliation: Department of Econometrics and Business Statistics, Monash University

keywords:
- data visualization
- statistical distributions
- time granularities
- calendar algebra
- periodicties
- grammar of graphics
- R

abstract: |
 Clustering elements based on behavior across time granularities 
bibliography: [bibliography.bib]
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
  \def\mod{~\text{mod}~}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 4
    fig_width: 6
    fig_caption: yes
    dev: "pdf" 
    keep_tex: yes
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
```

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk(here('script/smart_meter.R'))
```

# Introduction

@@

\noindent The massive amount of data generated in such projects could be overwhelming for analysis. Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers. Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data. Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors. In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior. Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors. The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis. One of the advantages of using our 
approach is that the technique is based on probability distributions instead of raw data. Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets. Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.

@@

# Preliminary exploratoration

## Electricity demand data

<!-- _Electricity use interval reading data_   -->
<!-- The data from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee). -->

### Getting data
The entire data is procured from CSIRO. A subset of this data is also available from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) is available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee). It consists of the following data sets. 
_1. CustomerData:_ 78720 customers with 62 variables about them 
_2. EUDMData:_ 300 billion half-hourly consumption level data  
_3. OffersData:_ Method of contact to customer to join SGSC customer trial, either door-to-door (D2D) or via Telesales  
_4. PEResponseData:_ Peak Events response customer wise  
_5. PETimesData:_ Peak Events time stamps

Only _CustomerData_ and _EUDMData_ are relevant for the clustering goals of this paper. _EUDMData_ contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total. `CustomerData` provides demographic data for 78,720 customers with information on their Local Government Area amongst others.

### Missing Data

Electricity usage for some customers may become unavailable due to power outage or not recording their usage properly, thus resulting in implicit missing values in the database. It is interesting to explore where missing-ness occurs or if there is a relationship between the underlying missing patterns. We use the R package `tsibble` to do this.

```{r, out.width="100%"}
include_graphics("figs/missing-data.png")
```

## LGA and weather

Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths
for each household. Since, general supply is available for only 13,735 customers, we will restrict ourselves to look at the LGA information for these customers only. We find that there are only 26 LGA that is covered for these customers.

_Weather data_  
This data is obtained through [Australian Government Bureau of Meteorology](BOM)(http://www.bom.gov.au/) and provides hourly data for nearest weather stations for all the LGAs

This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process. The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016. The data is at SA2 and LGA levels. However, some of the LGA in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs. Weather, notably temperature (and humidity) can be the main driver(s) for energy usage. In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use. Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au). Some weather stations have 30 minute (sometimes even smaller interval) weather data. Potentially, there could be lag effects of weather on energy usage which should be considered.

\noindent A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset. The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set. Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing. It looks like data is most missing before 2013 and for a particular date in 2014.

@@

```{r miss-data}
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

@@

```{r vismiss, eval = FALSE}
## transform all columns
customer_data <- customer_data %>% mutate_each(funs(empty_as_na))

vis_miss(customer_data, warn_large_data = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, size = 5))
```

```{r miss-var, eval = FALSE}
gg_miss_var(customer_data) + theme_bw() +
  theme(axis.text.y = element_text(size = 4))
```

```{r,eval=FALSE}
elec_ts <- read_rds("data/elec_ts.rds")
```

```{r raw-50, fig.height = 8, fig.cap="The raw data for 50 households are shown. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.", eval = FALSE}

@@

# Clustering methodology

@@

The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points. In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand. So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.

We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote

Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. Each household consists of a $J*K$ distributions one harmony. We compute the distributional difference
between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies. If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony. We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features. The features should also characterize probability distributions of different household.

<!-- grouping probability distributions across a harmony. This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data. The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities. The following are some of the advantages of our proposed method. -->

@@

```{r linear-cluster, eval = FALSE}
group1 <- all_data_cluster %>% 
  filter(groups == 1)

@@

## Cluster validation

**Internal cluster validation** uses the internal information of the clustering process to evaluate the goodness of a clustering structure without reference to external information. It can be also used for estimating the number of clusters and the appropriate clustering algorithm without any external data.

**External cluster validation** consists in comparing the results of a cluster analysis to an externally known result, such as externally provided class labels. It measures the extent to which cluster labels match externally supplied class labels. Since we know the “true” cluster number in advance, this approach is mainly used for selecting the right clustering algorithm for a specific data set.

**Relative cluster validation** evaluates the clustering structure by varying different parameter values for the same algorithm (e.g.,: varying the number of clusters k). It’s generally used for determining the optimal number of clusters.

Interval validation includes
_Compactness or cluster cohesion:_ Measures how close are the objects within the same cluster. A lower within-cluster variation is an indicator of a good compactness (i.e., a good clustering). The different indices for evaluating the compactness of clusters are base on distance measures such as the cluster-wise within average/median distances between observations.

_Separation:_ Measures how well-separated a cluster is from other clusters. The indices used as separation measures include:
distances between cluster centers
the pairwise minimum distances between objects in different clusters

_Connectivity:_ corresponds to what extent items are placed in the same cluster as their nearest neighbors in the data space. The connectivity has a value between 0 and infinity and should be minimized.

Generally most of the indices used for internal clustering validation combine compactness and separation measures.

@@

<!-- <Earlier write up> -->
<!-- ## Introduction -->

<!-- <!-- Description of data available --> -->
<!-- The Smart Grid, Smart City (SGSC) project, which rolled out Australia's first commercial-scale smart grid was implemented across eight local government areas in New South Wales (NSW). Data from more than 13,000 household electricity smart meters is obtained as part of that project. It provides half-hourly energy usage and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors. The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone. The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations. -->

<!-- The massive amount of data generated in such projects could be overwhelming for analysis. Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers. Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data. Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors. In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior. Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors. The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis. One of the advantages of using our  -->
<!-- approach is that the technique is based on probability distributions instead of raw data. Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets. Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data. -->

<!-- The following contributions are made through the following chapter: -->

<!--  * Present a cluster analysis of SGSC dataset to group households with similar periodic behavior -->
<!--  * Cluster validation by relating to external data -->

<!-- ## Introduction -->

<!-- In Australia, the federal government and fourteen partners, electricity distribution and transmission companies, technology companies, universities and CSIRO, invested AUD 490 million in the Smart Grid Smart City (SGSC) field trial. One of the largest commercial-scale smart grid technology assessment projects worldwide, SGSC is implemented across eight local government areas in New South Wales (NSW), covering 30,000 dwellings since 2009. -->
<!-- The SGSC database is publicly available through the Information Clearing House (ICH) (Ausgrid 2014). It is a massive database covering most physical aspects of the smart grid, as well as, aspects of users’ demographics. -->

<!-- Customer Sale and Half-Hour Power Consumption -->
<!-- and Generation Datasets -->

<!-- This is a big portion of the data from about 5900 households who accepted a -->
<!-- product out of about 24,000 offers made to the target population. Spatial resolution -->
<!-- down to anonymous individual dwellings [Customer Sales: Households: -->
<!-- Customer_Key] is also given based on {[Postcode], [Local Government Area], or -->
<!-- [Suburb Name]}. In order to make these datasets easily understandable, here we -->
<!-- present a new narrative approach to explain the available data fields and their -->
<!-- interdependencies in a brief yet logical and memorisable way as follows: -->

<!-- The [Half-Hour Consumption and Generation] data provide kWh readings of -->
<!-- the actual flow of energy. This includes the electricity supply imported from the -->
<!-- grid [General Supply] which is the main source of energy under any of the applied -->
<!-- tariffs, import during off-peak or for a control load [Off Peak], gross or net values of -->

<!-- domestic (often solar) electricity that is exported back to the grid [Generation]. The -->
<!-- [Peak Events] data field gives the details of every event [Event Key] including its -->
<!-- [Type], [Date], [Start Time], and [End Time]. Accordingly, the entire sequence of -->
<!-- user demand responses to any of the cost-reflective products is available in the -->
<!-- dataset to evaluate the project objectives as previously discussed. -->

<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings. For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home. These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state. Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers. Clustering households using only smart meter consumption data should provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households. -->

<!-- <!-- Description of data available --> -->
<!-- Data from more than 13,000 household electricity smart meters is obtained as part of the Smart Grid Smart City (SGSC) project (2010-2014). It provides one of the few linked sets of customer time of use (half-hour increments) and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors. The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone. This data set includes electricity use interval readings for each customer for every 30 minutes and demographic data for Australia, as well as detailed information on appliance use, climate, peak events; peak events response and retail and distributor product offers. The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations. -->

<!-- <!-- Background literature and shortcomings --> -->
<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics. The paper evaluates approximately 2100 unique peer-reviewed papers and presents three main findings related to clustering methods, data and cluster validation. None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data. The clustering is frequently applied directly to the raw data without investigating time series features like auto correlation and periodicity. The application of smart meter data to cluster electricity consumption using K-Means is ubiquitous. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. @Tureczek2018-ha has shown that the clever transformation of data before K-Means clustering can improve performance and enable K-Means. However, it does not explain the cluster composition by combining it with external data. We propose the following for the study of the Australian smart meter data. -->

<!-- Analysis for more than 13,000 household -->
<!-- electricity smart meters are included in this chapter. The following contributions are made through the following chapter: -->

<!-- - Presenting a cluster analysis of Australian residential electricity consumption data. -->

<!-- - Transformation and extraction of input data features enabling K-Means to account for other time series features in the clustering. This can easily be extended to include other data structures. -->

<!-- - Cluster validation by relating to external data -->

<!-- ## Preliminary exploratoration -->

<!-- This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations) and also the external data that needs to be utilized for validating the clustering process. -->

<!-- ## Common load clustering techniques of smart meter data -->

<!-- This section will discuss some literature on clustering techniques that were applied to smart meter customers in the past. -->

<!-- ## The proposed load-clustering technique -->

<!-- We employ a cluster analysis technique to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a cyclic granularity. This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data. The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities. Further, we elaborate on the effect of missing data and how it is handled in the clustering process. -->

<!-- ### Methodology, results, and discussions -->

<!-- This will contain the data structure and detailed methodology to be employed for the cluster analysis. -->

<!-- ### Distinction, repeatability, and robustness metrics -->

<!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. -->

<!-- <!-- ## Comparison with traditional load clustering methods --> -->

<!-- <!-- This section will cover the homogeneity and heterogeneity of the clusters, how distinct they are and summarize the main features of the cluster. --> -->

<!-- ## Combining findings with external data -->

<!-- We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation. We aim to do this by providing a practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions. Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.  -->
<!-- We will present how matching other data sources might help us in validating our clustering approach. -->

<!-- We will present how mapping the customers to their respective LGA's and reflect on the clusters obtained in the earlier section. -->

<!-- We will present how mapping the customers to their respective income and population level reflect on the clusters obtained in the earlier section. -->

<!-- ## Conclusion -->

<!-- This section will cover some drawback of this clustering method and potential extensions of this work. -->

<!-- There are many cluster analysis methods that can produce quite different clusterings on the same dataset. Cluster validation is about the evaluation of the quality of a clustering; “relative cluster validation” is about using such criteria to compare clusterings. This can be used to select one of a set of clusterings from different methods, or from the same method ran with different parameters such as different numbers of clusters. There are many cluster validation indexes in the literature. Most of them attempt to measure the overall quality of a clustering by a single number, but this can be inappropriate. There are various different characteristics of a clustering that can be relevant in practice, depending on the aim of clustering, such as low within-cluster distances and high between-cluster separation. In this paper, a number of validation criteria will be introduced that refer to different desirable characteristics of a clustering, and that characterise a clustering in a multidimensional way. In specific applications the user may be interested in some of these criteria rather than others. A focus of the paper is on methodology to standardise the different characteristics so that users can aggregate them in a suitable way specifying weights for the various criteria that are relevant in the clustering application at hand. -->

>>>>

---
title: "rough"
author: "Sayani Gupta"
date: "08/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
```

# Clustering Australian Smart Grid Smart City data for grouping periodically behaving customers

## Introduction

@@

The massive amount of data generated in such projects could be overwhelming for analysis. Electricity utilities can utilize the consumption patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers. Beyea (2010) has pointed out, there has been little discussion or exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data. Thus, there is a scope to investigate and analyze these data in various ways for a greater understanding of consumption patterns and how they correlate with other economic, physical or geographical factors. In this work, we are interested to see how we can utilize this dataset to group different customers with similar periodic behavior. Towards this goal, this chapter aims to: (a) describe the contents of the data set in SGSC database that we can utilize, and (b) propose a clustering algorithm to group customers with similar periodic behaviors. The distance metric introduced in Chapter 2 will be the inputs for this cluster analysis. One of the advantages of using our 
approach is that the technique is based on probability distributions instead of raw data. Many clustering approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets. Since the distance measure considered is based on differences in probability distribution of time series, it is likely to be less sensitive to missing or noisy data.

@@

## Preliminary exploratoration

This section familiarizes the 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations/number of customers/number of observations) and also the external data that needs to be utilized for validating the clustering process. The [ABS TableBuilder](https://www.abs.gov.au/websitedbs/censushome.nsf/home/tablebuilder) has census data from 2011 and 2016. The data is at SA2 and LGA levels. However, some of the LGAs in NSW changed between 2011 and 2016 and hence there would not be a one-to-one correspondence between the LGAs Weather, notably temperature (and humidity) can be the main driver(s) for energy usage.  In NSW many households have electric heaters so their use can impact winter energy use and air-conditioners can impact summer energy use. Relevant weather data could be obtained from the [Bureau of Meteorology](www.bom.gov.au). Some weather stations have 30 minute (sometimes even smaller interval) weather data. Potentially, there could be lag effects of weather on energy usage which should be considered.

A dataset of 100 SGSC homes has been used to lay out the structure to be used for analyzing the big dataset. The smaller dataset contains half-hourly kwh values form 2012 to 2014 and has asynchronous time series distributed evenly over the observation period (Figure \ref{fig:elec-raw}), similar to the bigger data set. Figure \ref{fig:count-gaps} can be used to interpret missingness in the data, where the customers are arranged from maximum to minimum missing. It looks like data is most missing before 2013 and for a particular date in 2014.

@@

```{r elec-raw, fig.height = 8, fig.cap="The raw data for 50 households are shown. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed."}

@@

```{r elec-gaps, warning=FALSE}

@@

```{r count-gaps, fig.height = 8, fig.cap="The missing data pattern for 100 households are shown. It looks like most missingness happens before 2013 and for a particular data in 2014."}

@@

The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_s(v)$ denotes the probability distribution from observation $s$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points. In our work, we are using $s$ as denoting a household and the underlying variable as the electricity demand. So $f_s(v)$ is the distribution of household $s$ and $v$ is electricity demand. Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. We still want the households in clusters, but each household consists of a collection of probability distributions. For  example, we would have $J*K$ distributions per household. So we need a distance between collections of univariate probability distributions. One possibility is to consider the normalized weighted pairwise distance as the difference between various households for the variable $(A, B)$. Let $wpd_{norm_{s,t}}(A,B)$ denotes the normalized pairwise distances between subjects/elements $s$ and $t$ for the harmony pair $(A, B)$. $wpd_{norm_{s,t}}(A,B)$ could be thought to be a dimensionality reduction procedure in time series data mining. This feature, which is considered as a measure to determine the strength of any periodic pattern are then inputs to a clustering algorithm that can then discern similarity and differences between the time series.

We employ a hierarchical cluster analysis algorithm to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a harmony. This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data. The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities. The following are some of the advantages of our proposed method.

@@

```{r linear-cluster}
group1 <- all_data_cluster %>% 
  filter(groups == 1)

@@

@@

```{r data_structure_clust, out.width = "100%", eval=FALSE}
knitr::include_graphics("figures/data_structure_clust.png")
```

<!-- - Study of Australian smart meter data -->

<!-- <!-- Motivation and background --> 
<!-- Electricity smart meter technology is increasingly being deployed in residential and commercial buildings. For example, in the state of Victoria, Australia, it is a state government policy that all households have a smart meter fitted (Victoria State Government, 2015), which has resulted in 97% of the approximately 2.1 million households in Victoria having a smart meter installed in their home. These smart meters collect energy usage information at half-hourly intervals, resulting in over 35 billion half-hourly observations per year across all households in the state. Clustering electricity consumption patterns can enable electricity utilities to develop targeted tariffs for individual groups alleviating the problem of volatility in production by capitalizing on the flexibility of consumers. Clustering households using only smart meter consumption data should provide value in a societal setting by combining the findings and external data like weather conditions, socio-economic or other demographic factors of those households. -->

<!-- <!-- Description of data available --> 
<!-- Data from more than 13,000 household electricity smart meters is obtained as part of the Smart Grid Smart City (SGSC) project (2010-2014). It provides one of the few linked sets of customer time of use (half-hour increments) and demographic data for Australia, as well as detailed information on appliance use, climate, retail and distributor product offers, and other related factors. The trials were based in Newcastle, New South Wales, but also covered areas in Sydney CBD, Newington, Ku-Ring-Gai, and the rural township of Scone. This data set includes electricity use interval readings for each customer for every 30 minutes and demographic data for Australia, as well as detailed information on appliance use, climate, peak events; peak events response and retail and distributor product offers. The load time series is asynchronous as it is observed for these households for unequal time lengths and consists of missing observations. -->

<!-- <!-- Background literature and shortcomings --> 
<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics. The paper evaluates approximately 2100 unique peer-reviewed papers and presents three main findings related to clustering methods, data and cluster validation. None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data. The clustering is frequently applied directly to the raw data without investigating time series features like auto correlation and periodicity. The application of smart meter data to cluster electricity consumption using K-Means is ubiquitous. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. @Tureczek2018-ha has shown that the clever transformation of data before K-Means clustering can improve performance and enable K-Means. However, it does not explain the cluster composition by combining it with external data. -->

<!-- We propose the following for the study of the Australian smart meter data: -->

<!--  - *Preliminary exploratory visualization and summarisation*: Be familiar with an overall view of these 13,000 SGSC households through visualization and provide a detailed layout of data structures (like missing observations) and also the external data that needs to be utilized in the clustering phase. -->

<!--  - *Cluster analysis*: We employ a cluster analysis technique to obtain pockets of households showing similar periodic behavior by grouping probability distributions across a cyclic granularity. This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data. The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities. Further, we elaborate on the effect of missing data and how it is handled in the clustering process. We will also determine if any identified clusters or patterns are indeed statistically meaningful in the sense that they actually exist and are not a random allocation. We aim to do this by providing a practical explanation of the formed clusters, visualizing how they relate to any weather, socio-economic or geographical conditions. Furthermore, we ensure that a cross-validation algorithm is employed leading to robust clusters.  -->

<!-- NOT USED IN MID_CAN -->

<!-- ## Clustering households with similar periodic behavior -->

<!-- The pool of all customers’ load time series will have different periods of time with overlaps. Any feature based clustering approach would trim the load curves to obtain equal length and synchronized date and time for all curves. These approaches are limited by the type of noisy, patchy, and unequal time-series common in residential data sets (@). In this project, we explore a clustering strategy to alleviate these limitations by looking at the distribution of load time series across different non-linear time granularities that can be readily clustered. Clustering could be employed for distribution of any harmonies (or all non-linear granularities) taken together to form clusters with similar periodic behavior across these harmonies (or all non-linear granularities). @ suggested carrying out cross validation by bootstrapping to form same clusters of homes on randomly selected days within a certain period. The strategy employed here when complimented with some cross validation would be a way to form same clusters of homes who show similar periodic behavior on randomly selected harmonies within a certain period. -->

# Timeline
